Analyzing Fairness: Race, Recidivism, and the COMPAS Algorithm

### 2022-04-21 final_metrics_analysis.Rmd

```{python exploratory race bar plot}
# barplot of race
sns.countplot(x="race", data=df_viz, palette=workshop_palette)
plt.title('Race Distribution (White and Black individuals only)')
```

@@

```{python post-processing equal odds needs fix} 
privileged_groups, unprivileged_groups = get_attributes(train_BLD, selected_attr= 'race')
ep = EqOddsPostprocessing(unprivileged_groups= unprivileged_groups, privileged_groups= privileged_groups, seed = 42)
#eo = PostProcessingMeta(estimator=lr, postprocessor=ep, random_state=1234567)
EOPP = ep.fit(train_BLD, test_BLD)
y_pred_EO = ep.predict(test_BLD)
#y_proba_EO = ep.predict_(test_BLD)

acc_EO = accuracy_score(y_test, y_pred_EO)
```

>>>>

```{python crosstable of recitivism by race}
crosstable=pd.crosstab(df_viz['race'],df_viz['two_year_recid'])
crosstable
```

```{python exploratory recidivism mosaic plot}
props={}
props[('Caucasian','Recidivated')]={'facecolor':'red', 'edgecolor':'white'}
props[('Caucasian','Survived')]={'facecolor':'red', 'edgecolor':'white'}
props[('African-American','Recidivated')]={'facecolor':'xkcd:aqua','edgecolor':'white'}
props[('African-American','Survived')]=        {'facecolor':'xkcd:aqua','edgecolor':'white'}
labelizer=lambda k:{('Caucasian','Recidivated'):822,('African-American','Recidivated'):1661,('Caucasian','Survived'):1278,('African-American','Survived'):1512}[k]
mosaic(df_viz,['race','two_year_recid'],labelizer=labelizer,properties=props)
plt.show()
```

A higher percentage of white defendants (60.86%) than black defendants (47.65%) did not commit another crime within two years. This means that 39.14% of white defendants did recidivate within two years compared to 52.35% of black defendants.

```{python exploratory race bar plot}
# barplot of race
sns.countplot(x="race", data=df_viz, palette=workshop_palette)
plt.title('Race Distribution (White and Black individuals only)')
plt.show()
```

@@

```{python post-processing equal odds needs fix} 
#privileged_groups, unprivileged_groups = get_attributes(train_BLD, selected_attr= 'race')
#ep = EqOddsPostprocessing(unprivileged_groups= unprivileged_groups, privileged_groups= privileged_groups, seed = 42)
#eo = PostProcessingMeta(estimator=lr, postprocessor=ep, random_state=1234567)
#EOPP = ep.fit(train_BLD, test_BLD)
#y_pred_EO = ep.predict(test_BLD)
#y_proba_EO = ep.predict_(test_BLD)

#acc_EO = accuracy_score(y_test, y_pred_EO)
```
